---
title: "Download and install CONFETTI from GitHub"
author: "Felix May"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Download and install CONFETTI from GitHub}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installing the package from GitHub

To install and build the package from source directly from GitHub you can use the `devtools` package. In advance you need to install the package `Rcpp` as well as a  an appropriate C++ compiler. Under Windows this means you need Rtools34from CRAN. If you have these tools than installing the package is very easy:

```{r, message=F}
library(Rcpp)
library(devtools)
install_github("FelixMay/confettiRbasic", build_vignettes = T)
library(confettiRbasic)
```

# Getting help

There is just one main function onc the package

```{r, message=FALSE}
?confetti.run
```

However, there are also vignettes that introduce the use of the model in detail

```{r, message=F}
help(package = "confettiRbasic")
browseVignettes("confettiRbasic")
```


# Downloading the source code from GitHub

You can also get the source code from GitHub

[https://github.com/FelixMay/confettiRbasic](https://github.com/FelixMay/confettiRbasic)

Use the button **"Clone or download"** do download the package source code as zip-file.

However, it is highly recommended to link you copy to the main repository by forking
the repo to your own account as explained in the repo's page.
